---
title: "Clustering"
author: "Francis"
date: "October 18, 2018"
output: html_document
---

In this section we will use the Iris data set to experiment with different clustering techniques.

Let's load some data
```{r}
library(datasets)
head(iris)
data<-iris
```

Exploratory data analysis
```{r}
library(ggplot2)
ggplot(iris, aes(Petal.Length, Petal.Width, color = Species)) + geom_point()
```

Our first clustering technique will be k-means clustering.  K-means clustering is an unsupervised learning method that associates data based on similarity.  Here we develop our clusters.
```{r}
set.seed(20)
Iriscluster <- kmeans(iris[, 3:4], 3, nstart = 20)
Iriscluster
```

Let's examine the effectiveness of our clusters.
```{r}
table(Iriscluster$cluster, data$Species)
```

We can also view a graphical representation of our clusters.
```{r}
Iriscluster$cluster <- as.factor(irisCluster$cluster)
ggplot(iris, aes(Petal.Length, Petal.Width, color = Iriscluster$cluster)) + geom_point()+geom_polygon()
```

