---
title: "FBI"
author: "Francis"
date: "August 26, 2018"
output: html_document
---

Let's load some data 
The data we received was in an uneditable excel file.  To workaround we created a .txt file and deleted the column headers.
```{r}
fbi<-read.table("C:/Users/Owner/Desktop/fbi.txt", header = F)
head(fbi)
```
Now we will put back the data element headers

```{r}

```

Let's subset some data
Year~population
```{r}
pop<-fbi[c(1,2)]
names(pop)<-c("year","population")
head(pop)
r<-fbi[c(1,3)]
names(r)<-c("year","rate")
head(r)

```

Let's explore ggplot2
```{r}
library(ggplot2)
ggplot()+geom_point(data=pop, aes(year,population))+geom_point(data = r,aes(year,rate))


```

Jerry: make more interesting graphs with other variables.


Let's start some interpolation and extrapolation

Interpolation: approximating values in between different data points (estimation)
Here we are attempting to approximate population by month.  n= 12 Months * 20 Years
```{r}
approx(pop$year, pop$population, method = "linear", n=228)
popMonth<-approx(pop$year, pop$population, method = "linear", n=240)
```
Next we will refactor the decimal dates into months 


Extrapolation: approximating values outside of different data points (prediction)
```{r}

```