---
title: "OutlierAnalysis"
author: "Francis"
date: "May 30, 2018"
output:
  word_document: default
  pdf_document: default
  html_document: default
---
Outlier Treatment
To investigate the importance of removing outliers we will add outliers to the cars dataset.
Adding outliers
```{r}
cars1 <- cars[1:30, ]  # original data
cars_outliers <- data.frame(speed=c(19,19,20,20,20), dist=c(190, 186, 210, 220, 218))  # introduce outliers.
cars2 <- rbind(cars1, cars_outliers)  # data with outliers.
```
Plot data with outliers.  (use ggplot)
```{r}
par(mfrow=c(1, 2)) #parse into 2 columns
plot(cars2$speed, cars2$dist, xlim=c(0, 28), ylim=c(0, 230), main="With Outliers", xlab="speed", ylab="dist", pch="*", col="red", cex=2)
abline(lm(dist ~ speed, data=cars2), col="blue", lwd=3, lty=2)
```
Plot without outliers
```{r}
plot(cars1$speed, cars1$dist, xlim=c(0, 28), ylim=c(0, 230), main="Outliers removed \n A much better fit!", xlab="speed", ylab="dist", pch="*", col="red", cex=2)
abline(lm(dist ~ speed, data=cars1), col="blue", lwd=3, lty=2)
```
Detecting outliers
1. Univariate approach
For a continuous variable, an outlier is considered to be an observation that lies outside 1.5*IQR or, the inter quartile range.
```{r}
url <- "http://rstatistics.net/wp-content/uploads/2015/09/ozone.csv"
inputData <- read.csv(url)  # import data
outlier_values <- boxplot.stats(inputData$pressure_height)$out  # outlier values.
boxplot(inputData$pressure_height, main="Pressure Height", boxwex=0.1)
mtext(paste("Outliers: ", paste(outlier_values, collapse=", ")), cex=0.6)
```
2. Bivariate approach
visualize using box-plots of x and y for categorical X's
```{r}
url <- "http://rstatistics.net/wp-content/uploads/2015/09/ozone.csv"
ozone <- read.csv(url)
# For categorical variable
boxplot(ozone_reading ~ Month, data=ozone, main="Ozone reading across months")  # clear pattern is noticeable.
boxplot(ozone_reading ~ Day_of_week, data=ozone, main="Ozone reading for days of week")

# For continuous variable (convert to categorical if needed.)
boxplot(ozone_reading ~ pressure_height, data=ozone, main="Boxplot for Pressure height (continuos var) vs Ozone")
boxplot(ozone_reading ~ cut(pressure_height, pretty(inputData$pressure_height)), data=ozone, main="Boxplot for Pressure height (categorial) vs Ozone", cex.axis=0.5)
```
3. Multivariate approach
Cook's distance:  measures how much each observation impacts the fitted values
```{r}
mod <- lm(ozone_reading ~ .,data = ozone)
cooksd <- cooks.distance(mod)
```
Observations with a cook's distance greater than 4 times the mean are influential.
```{r}
plot(cooksd, pch="*", cex=2, main="Influential Obs by Cooks distance")  # plot cook's distance
abline(h = 4*mean(cooksd, na.rm=T), col="red")  # add cutoff line
text(x=1:length(cooksd)+1, y=cooksd, labels=ifelse(cooksd>4*mean(cooksd, na.rm=T),names(cooksd),""), col="red")  # add labels
```
identify infuential obs
```{r}
influential <- as.numeric(names(cooksd)[(cooksd > 4*mean(cooksd, na.rm=T))])  # influential row numbers
head(ozone[influential, ])  # influential observations.
```

Outliers test
```{r}
library("car")
library("outliers")
car::outlierTest(mod)
```
Outliers Package
outliers(): gets the most extreme observation from the mean.
```{r}
set.seed(1234)
y=rnorm(100)
outlier(y)
#> [1] 2.548991
outlier(y,opposite=TRUE)
#> [1] -2.345698
dim(y) <- c(20,5)  # convert it to a matrix
outlier(y)
outlier(y,opposite=T)
```
scores(): computes the normalized scores and finds observations that lie outside a given percentile
```{r}
set.seed(1000)
x = rnorm(10)
scores(x)
scores(x, type = "chisq")
scores(x, type = "t")
scores(x, type = "chisq", prob = 0.9)
scores(x, type="chisq", prob=0.95)  # beyond 95th %ile
scores(x, type="z", prob=0.95)  # beyond 95th %ile based on z-scores
scores(x, type="t", prob=0.95)  # beyond 95th %ile based on t-scores
```
Treating the outliers 
1. Imputation
same as above
2. Capping replacing values that lie outside 1.5*IQR with the value of the 5th percentile if the outlier is bellow the lower limit and the value of the 95th percentile if the outlier lies above the upper limit.
```{r}
x <- ozone$pressure_height
qnt <- quantile(x, probs=c(.25, .75), na.rm = T)
caps <- quantile(x, probs=c(.05, .95), na.rm = T)
H <- 1.5 * IQR(x, na.rm = T)
x[x < (qnt[1] - H)] <- caps[1]
x[x > (qnt[2] + H)] <- caps[2]
```
3. Prediction
See above
